---
title: "Corporate Governance"
author: "Arunendu (10)"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
# ARUNENDU BIKAS CHAKI 251260250010

df <- read.csv(choose.files())
View(df)
```
```{r}
 install.packages("readr")
 install.packages("dplyr")
 install.packages("lmtest")
 install.packages("car")
```
```{r}
df <- read.csv('corporate_governance.csv')
print("Clean Data:")
```
```{r create-dummies}
 df <- read.csv('corporate_governance.csv', stringsAsFactors = FALSE)
print(colnames(df))
df_clean <- df %>%
  rename(
    Firm = FIRM,
    ROA_str = ROA..Approx.,
    Board_Size = BoardSize,
    Indep_Percent_str = Independent.Directors.,
    Audit_Meetings = Audit.Committee.Meetings
  ) %>%
  mutate(
    ROA = as.numeric(gsub("%", "", ROA_str)),
    Indep_Percent = as.numeric(gsub("%", "", Indep_Percent_str))
  ) %>%
  select(Firm, ROA, Board_Size, Indep_Percent, Audit_Meetings) %>%
   na.omit()
print("Clean Data:")
head(df_clean)
```

```{r create-dummies}
df_clean <- df_clean %>%
  mutate(
    High_Indep = ifelse(Indep_Percent >= 50, 1, 0),
    Many_Meetings = ifelse(Audit_Meetings > 4, 1, 0)
  )

print("Data with Dummies:")
head(df_clean)
```
Now we will run our 6 regression models to test our theories. Our dependent variable (what we want to explain) is `ROA`.

MODEL 1
ROA vs. Board Size
This model asks a simple question: **"Does a bigger board relate to higher or lower ROA?"**
```{r }
m1 <- lm(ROA ~ Board_Size, data = df_clean)
summary(m1) 
```
Interpretation:
The intercept (7.65) suggests that when the board size is zero (a hypothetical case), the firm’s ROA would be around 7.65%.
The coefficient of Board_Size (−0.0994) indicates a slight negative relationship — for each additional board member, the ROA decreases by about 0.1 percentage points on average.

Statistical Significance:
The p-value (0.832) is much greater than 0.05, meaning the effect of board size on ROA is not statistically significant.
The R-squared (0.0016) means the model explains only 0.16% of the variation in ROA — essentially no explanatory power.

Model 2: 
ROA vs. Percent of Independent Directors
This model asks: "Does having more independent directors relate to higher or lower ROA?"
```{r}
m2 <- lm(ROA ~ Indep_Percent, data = df_clean)
summary(m2)
```
Interpretation:
The intercept (10.97) indicates that if the percentage of independent directors were zero (hypothetically), the firm’s ROA would be about 10.97%.
The coefficient for Indep_Percent (−0.0855) shows a small negative relationship — as the proportion of independent directors increases by 1%, the ROA decreases by about 0.086 percentage points on average.

Statistical Significance:
The p-value (0.714) is far above 0.05, so this relationship is not statistically significant.
The R-squared (0.0049) means the model explains less than 1% of the variation in ROA, indicating very poor explanatory power.

Model 3: 
The Full Model
This model asks:"What is the effect of all three variables (Board Size, Independence, and Meetings) on ROA at the same time?"
```{r}
m3 <- lm(ROA ~ Board_Size + Indep_Percent + Audit_Meetings, data = df_clean)
summary(m3)
```
Interpretation: This is a stronger model. It tells you the effect of `Board_Size` while holding the other two variables constant.
Intercept (3.96):
When board size, independence percentage, and audit meetings are zero (hypothetically), ROA would be around 3.96%.
Board_Size (0.178, p = 0.721):
A small positive but insignificant effect — changes in board size do not significantly affect ROA.
Indep_Percent (0.108, p = 0.672):
A positive but insignificant relationship — a higher percentage of independent directors does not have a statistically significant impact on profitability.
Audit_Meetings (−0.852, p = 0.035):
This coefficient is negative and statistically significant at the 5% level.
→ It suggests that more frequent audit committee meetings are associated with lower ROA, possibly due to increased monitoring during periods of weaker performance.

Model Fit:
R² = 0.1685: The model explains about 16.9% of the variation in ROA.
Adjusted R² = 0.0726: After accounting for the number of predictors, only about 7% of variation is explained.
F-statistic (p = 0.180): The overall model is not statistically significant.


Model 4 :- 
The "Curvy" Model (Quadratic)
This model tests a theory: "Is there an 'optimal' board size?" Maybe ROA goes up as boards get bigger, but then goes down if they get too big. This tests for an "inverted U-shape".
```{r}
m4 <- lm(ROA ~ Board_Size + I(Board_Size^2) + Indep_Percent + Audit_Meetings, data = df_clean)
summary(m4)
```
Interpretation: The model explores how board size (including its squared term), the percentage of independent directors, and the number of audit meetings affect a firm’s profitability (ROA).
The coefficients of Board_Size (−5.93) and Board_Size² (0.27) suggest a U-shaped relationship between board size and ROA — profitability initially falls as board size increases, but beyond a certain point, larger boards may improve performance. However, both effects are not statistically significant (p > 0.1).
Indep_Percent (0.11, p = 0.66) shows a weak and insignificant positive relationship with ROA.
Audit_Meetings (−0.71, p = 0.075) is negative and marginally significant, implying that firms holding more audit meetings tend to have slightly lower profitability, possibly due to greater scrutiny during weaker financial performance.
The model explains about 24.6% of the variation in ROA (R² = 0.2456), but the overall model is not statistically significant (p = 0.12).

Model 5:  
The Dummy Variable Model
This model asks: "Does simply meeting the 50% independence threshold matter, rather than the exact percentage?"
```{r}
m5 <- lm(ROA ~ Board_Size + High_Indep + Audit_Meetings, data = df_clean)
summary(m5)
```
Interpretation:
Intercept (13.31, p = 0.064):
When all other variables are zero, the expected ROA is around 13.3%, though this is only marginally significant.
Board_Size (−0.079, p = 0.878):
The effect of board size on ROA is negative but statistically insignificant, suggesting that changes in board size do not meaningfully influence profitability.
High_Indep (−1.88, p = 0.513):
Firms with at least 50% independent directors have, on average, a 1.88 percentage point lower ROA, but this result is not statistically significant.
Audit_Meetings (−0.758, p = 0.041):
The number of audit meetings shows a negative and statistically significant relationship with ROA at the 5% level.
→ This indicates that firms holding more audit meetings tend to have lower profitability, possibly because frequent meetings occur when performance issues or risks arise.

Model Fit:
R² = 0.1766: The model explains about 17.7% of the variation in ROA.
Adjusted R² = 0.0816: After adjusting for predictors, only about 8% of variation is explained.
F-statistic (p = 0.161): The overall model is not statistically significant.

Model 6: 
The Interaction Model
This is the most advanced model. It asks: "Does the effect of independence depend on*how many meetings are held?" For example, maybe independence only helps if the committee also meets a lot.

```{r model-6}
m6 <- lm(ROA ~ Board_Size + Indep_Percent * Many_Meetings, data = df_clean)
summary(m6)
```
Interpretation:
Intercept (33.33, p = 0.252):
Represents the estimated ROA when all predictors are zero — mainly a baseline value, not practically meaningful.
Board_Size (−0.27, p = 0.609):
Shows a small negative but insignificant effect — board size does not significantly impact ROA.
Indep_Percent (−0.46, p = 0.380):
Indicates a negative but statistically insignificant relationship — a higher percentage of independent directors alone is not linked to higher profitability.
Many_Meetings (−42.92, p = 0.163):
Firms with more than 4 audit meetings tend to have a lower ROA, though the effect is not significant.
Interaction (Indep_Percent × Many_Meetings = 0.79, p = 0.194):
The positive interaction term suggests that when audit meetings are frequent, having a higher percentage of independent directors slightly improves ROA, but this effect is not statistically significant.

Model Fit:
R² = 0.138: About 13.8% of the variation in ROA is explained by the model.
Adjusted R² ≈ 0: After adjusting for the number of predictors, the model explains almost no variation.
F-statistic (p = 0.425): The overall model is not statistically significant.
```{r}
library(car)
library(lmtest)
vif(m3)
```
All VIF values (1.23–1.34) are well below 5, indicating no multicollinearity among the independent variables.
Hence, the model’s estimates are statistically stable and reliable.
```{r}
bptest(m3)
```

The Breusch-Pagan test checks for heteroskedasticity (unequal variance of residuals).
Here, BP = 3.05 and p-value = 0.3833, which is greater than 0.05, indicating no significant heteroskedasticity.
```{r}
dwtest(m3)
```
The Durbin-Watson test checks for autocorrelation (correlation among residuals).
Here, DW = 2.62 with a p-value = 0.9478, which is much greater than 0.05, indicating no positive autocorrelation.


SUMMARY 
Overall, the models indicate that:
Corporate governance variables (board size, independence, and audit meetings) have limited explanatory power for firm profitability.
Only audit committee activity occasionally shows a meaningful (negative) association.
The low R² values across all models suggest that other firm-level or external factors likely play a much larger role in determining profitability (ROA).